\MakeUppercase{Linear versus centred chromatic numbers}\thanks{This research was partly funded by NSERC.}

### 2024-04-05 lin-vs-cen.tex
Pffft

\documentclass{patmorin}
\listfiles
\usepackage{pat}
\usepackage{paralist}
\usepackage{dsfont}  % for \mathds{A}
\usepackage[OT1]{fontenc}
\usepackage[utf8x]{inputenc}
\usepackage{skull}
\usepackage{paralist}
\usepackage{graphicx}
\usepackage[noend]{algorithmic}
\usepackage{bbm}  % needed for \mathbbm{1}
\usepackage{listings}
\usepackage{thmtools}
\usepackage{thm-restate}

\newtheorem{zeb}[theorem]{Lemma}

\usepackage[normalem]{ulem}
\usepackage{cancel}
%\usepackage{enumitem}

@@

\begin{restatable}{zeb}{blechblech}\label{pick_up_two}
  Let $G$ be an $a\times a$ pseudogrid with $a\ge 5$,  let $s,v,w,t$ be vertices of $G$ with $s$ in column $1$ of $G$, $v,w\in \interior_1(G)$, and $t$ in column $a$ of $G$.  Then $G$ contains a path $P$ with endpoints $s$ and $t$ that contains $v$ and $w$.
\end{restatable}

@@

We are also grateful to Omer Angel for pointing out that the problem on a $k\times k$ grid (as opposed to pseudogrid) has a simple solution: Partition the grid into $2\times 2$ grids and repeatedly delete a subgrid if it contains a vertex whose colour is unique.  If this causes the grid to become disconnected, keep only the largest remaining component and continue.  This process deletes at most $4\epsilon k$ vertices.  For $\epsilon < 1/4$, this implies that the graph that remains at the end of thie process has at least $k^2/2$ vertices.  In particular, the graph that remains is non-empty and every colour that appears in this graph is used on at least two vertices.  Finally, this graph has a Hamiltonian path (which can be constructed by traversing the boundary of  a spanning-tree of the dual graph).

@@

\blechblech

>>>>

\documentclass{patmorin}
\listfiles
\usepackage{pat}
\usepackage{paralist}
\usepackage[OT1]{fontenc}
\usepackage[utf8x]{inputenc}
\usepackage{paralist}
\usepackage{bbm}  % needed for \mathbbm{1}

@@

\begin{lem}[restate=pickuptwo,label=pick_up_two]
  Let $G$ be an $a\times a$ pseudogrid with $a\ge 5$,  let $s,v,w,t$ be vertices of $G$ with $s$ in column $1$ of $G$, $v,w\in \interior_1(G)$, and $t$ in column $a$ of $G$.  Then $G$ contains a path $P$ with endpoints $s$ and $t$ that contains $v$ and $w$.
\end{lem}

@@

We are also grateful to Omer Angel for pointing out that the problem on a $k\times k$ grid (as opposed to pseudogrid) has a simple solution, illustrated in \cref{omer}: Partition the grid into $2\times 2$ grids and repeatedly delete a subgrid if it contains a vertex whose colour is unique.  If this causes the grid to become disconnected, keep only the largest remaining component and continue.  This process deletes at most $4\epsilon k$ vertices.  For $\epsilon < 1/4$, this implies that the graph that remains at the end of thie process has at least $k^2/2$ vertices.  In particular, the graph that remains is non-empty and every colour that appears in this graph is used on at least two vertices.  Finally, this graph has a Hamiltonian path (which can be constructed by traversing the boundary of a certain tree contained in the dual graph).  Unfortunately, this solution breaks down immediately for pseudogrids since a pseudogrid is not necessarily Hamiltonian, even before removing some of its vertices and edges.

\begin{figure}
  \centering
  \begin{tabular}{cccc}
    \includegraphics[page=1]{figs/omer} &
    \includegraphics[page=2]{figs/omer} &
    % \includegraphics[page=3]{figs/omer} &
    \includegraphics[page=4]{figs/omer}
  \end{tabular}
  \caption{A simpler solution for colourings of the grid. The (orange) tree used to create the (purple) Hamiltonian path is created so that its edges and leaves use only odd-numbered rows and columns of the dual graph.}
  \label{omer}
\end{figure}

@@

\pickuptwo*
